Operationally-Safe Peer-to-Peer Energy Trading in Distribution Grids: A Game-Theoretic Market-Clearing Mechanism

### 2022-01-11 papers/01 Reply round 1/response_TSG.tex

\query[R1-5]{The individual variables, coupled relationships for these energy roles, and the corresponding dual variables should be explained specifically, along with the whole framework structure for equilibrium. The rationalities of the framework is not clear. At least, the current presentation is hard to follow.  }
\ans{
\green{
In the revised paper, (i) we provided a detailed description of each individual variable along with an explanation of its function in the algorithm; (ii) we expanded the description of the algorithm; and (iii) we included a paragraph to present the framework structure for equilibrium considered. Please, refer to the margin notes (R1-5) in Section III for the relative changes and additions.
}
}

\query[R1-6]{The P2P trading prices are given parameters, which are quite unreasonable. In fact, the P2P prices are the key variable in most literature.}
\ans{In this manuscript, we want to consider the most general problem formulation of peer-to-peer markets available in the literature. Following \cite{sorin2019,lecadre2020,baroche2019b}, a linear trading cost function is introduced in the problem formulation for the purpose of product differentiation or preference prices. %Furthermore, \cite{baroche2019} also adds a trading tariff or transaction cost encoded as a 1-norm cost function. 
	Furthermore, we may set $c_{\{i,j\}}^{\mathrm{tr}}=0$, and thus only consider the dual variable associated with the reciprocity constraints (Eq. \red{(8b)} in the revised manuscript) as the shadow price of bilateral trading \cite[Section 2.4]{lecadre2020}. Indeed the proposed algorithm can handle any case where $c_{\{i,j\}}^{\mathrm{tr}} \geq 0$, for all $\{i,j\} \in \mc E$.}

 \query[R1-7]{The algorithm requires too many iterations, which leads to poor feasibility in real applications, especially in distributed environment. }
 \ans{\comment{For day-ahead or intra-day markets, our numerical experience shows the algorithm is tractable.}
\green{I have to answer}. 
 }

@@

\textit{``$\dots$ remain satisfied.} \edit{\textit{In addition, instead of considering linear power flow equations in \red{(14)}, a nonlinear convex relaxation, such as a second order cone or semi-definite programming as discussed in \cite[Sect. II-A]{molzahn2017} can be considered since it still satisfies Assumption 1. In this case, only the definition of $\mc U_{N+1}$ differs from the current formulation.
''}
}

\query[R3-3]{As well pointed by the authors, this is a semi-decentralized approach where a coordinator is needed. Please, provide some discussion on how this is different with community-based local markets.   }
\ans{As discussed in \cite[Sect. 3.2]{sousa2019}, in community-based local markets, each prosumer  does not know with which other prosumers it trades energy. The power balance of the whole network (all prosumers) is considered. Our setup considers full P2P markets where each prosumer knows its trading partners and, thus, negotiates directly with them. The coordinator in our setup plays a role in handling the physical constraints of the network as well as the aggregated power bought from the main grid.}

\query[R3-4]{ The proposed algorithm relies on several dual variables. Please, elaborate on the meaning of this dual variables and their interpretation as economic outputs of the clearing process. }
%<<<<<<< HEAD
\ans{\green{
We have included a paragraph in Section III B in which we list all primal, dual, and auxiliary variables of the main algorithm, along with an explanation of their scope, and economical and system-theoretic interpretations.
}}
%=======
%\ans{We introduce dual variables for the purpose of problem decomposition. Particularly, these dual variables are associated with the coupling constraints that define $\mc C$ in \red{(9)}. Some of these dual variables, i.e., particularly those that are associated with bus power balance constraints \red{(13)} and reciprocity constraints \red{(8b)}, can be intrepreted as  nodal prices and bilateral trading prices, respectively, as discussed in \cite[Section 2.4]{lecadre2020}. }
%>>>>>>> fa2f889457721f5328b34fcf1ccafeda7a6ee6a0

@@

\newpage
\section*{Answers to Reviewer 4}
\textit{This paper presents a mathematical model for a P2P energy market formulated as a generalized game. They introduce in the game a new agent, a Distribution Network Operator (DNO), who is responsible of ensure the operational constraints of the market. Further, they propose a semi-decentralized algorithm to solve the problem, ensuring convergence to a Nash equilibrium. Finally, a numerical study is used to show the relevance of the market framework and the scalability of the algorithm.}

\query[R4-1]{In my opinion, the paper is clearly written, with a detailed explanation of the mathematical aspects, and introduces some relevant results. However, some questions need to be addressed in order to improve the quality of the work.  }
\ans{We thank the Reviewer for the positive comment.}

\textit{Comments and suggestions: }

\query[R4-2]{The authors state that due to the presence of coupling objective and constraints in the formulation, this implies the inapplicability of other decentralized algorithms employed in the literature (lines 54-57, page 2). This is not strictly true, since it is straightforward to reformulate the problem to decompose the objective function, just by replicating the variables avoiding the decomposition, and introduce new complicating constraints to coordinate these variables. Reformulate this sentence. }
\ans{\green{
We respectfully point out that the original game and the ``decomposed" one, obtained by including local slack variables and coupling constraints to coordinate these variables have, in general, different equilibria, see the example at the end of the answer. Therefore, social-welfare optimization methods (e.g., ADMM) cannot be straightforwardly used to solve our game-theoretic problem (which is a variational inequality). This is not surprising since v-GNEs and social optima, i.e., the minimizers of the social welfare, are essentially different configurations and coincide only in special cases. \\
In the revised paper, we rephrased the sentence avoiding strong claims of inapplicability, as follows:
}}

\textit{ \edit{``Nevertheless, none of these works considers both coupled objectives and constraints.''}}

\green{
\textit{Example.} Consider the 2-player aggregative game:
\begin{align*} 
\text{player }1: \quad \min_{u_1 \in \mathbb R}J_1(u_1,u_2) = (u_1+u_2 + 1)^\top u_1,\\
\text{player }2: \quad \min_{u_2 \in \mathbb R} J_2(u_2,u_1) = (u_1+u_2 + 1)^\top u_2.
\end{align*}
Its Nash equilibria can be found in closed form by solving the system of equations obtained by imposing the pseudo-gradient mapping of the game to zero, i.e.,
\begin{equation*}
\left[
\begin{array}{l}
\nabla_{u_1} J_1(u_1,u_2)\\
\nabla_{u_2} J_2(u_1,u_2)
\end{array}
\right]
=
\left[
\begin{array}{l}
2u_1+u_2+1\\
u_1+2u_2+1
\end{array}
\right]
\overset{!}{=}
\left[
\begin{array}{l}
0\\
0
\end{array}
\right],
\end{equation*} 
which yields a unique NE solution: $(u_1^*,u_2^*) = (-1/3,-1/3) $.\\
Now, consider the generalized game
\begin{align*} \
\text{player }1: \quad 
\left\{
\begin{array}{r l}
\displaystyle
\min_{u_1,\sigma_1 \in \mathbb R}& J_1(u_1,\sigma_1) = (u_1+\sigma_1 + 1)^\top u_1,\\
\text{s.t. } & \sigma_1 - u_2 = 0\\
            &  \sigma_2 - u_1 = 0
\end{array}
\right.
\\
\text{player }2: \quad 
\left\{
\begin{array}{r l}
\displaystyle
\min_{u_2,\sigma_2 \in \mathbb R}& J_2(u_2,\sigma_2) = (\sigma_2+u_2 + 1)^\top u_2,\\
\text{s.t. } & \sigma_1 - u_2 = 0\\
            &  \sigma_2 - u_1 = 0
\end{array}
\right.
\end{align*} 
obtained by decomposing the objective functions using the local slack variables $\sigma_1,\sigma_2$ and the extra coupling constraints $\sigma_1 - u_2 = 0, \, \sigma_2 - u_1$. Let us define the Lagrangian functions
\begin{align*}
\text{player }1: \quad \mc L_1(u_1,u_2,\sigma_1,\sigma_2,\lambda_{1,1},\lambda_{1,2}) &= (u_1+\sigma_1 + 1)^\top u_1 + \lambda_{1,1}^\top(\sigma_1 - u_2) + \lambda_{1,2}^\top(\sigma_2 - u_1),\\
\text{player }2: \quad \mc L_2(u_2,u_1,\sigma_2,\sigma_1,\lambda_{2,1},\lambda_{2,2}) &= (u_2+\sigma_2 + 1)^\top u_2 + \lambda_{2,1}^\top(\sigma_1 - u_2) + \lambda_{2,2}^\top(\sigma_2 - u_1),
\end{align*}
The variational GNEs of this modified game can be obtained in closed form by solving the system of equations obtained by stacking the coupled KKT systems of player 1 and 2, with equal dual variables ($\lambda_{1,1}=\lambda_{2,1} := \lambda_1$ and $\lambda_{2,1}=\lambda_{2,2} :=\lambda_2$) \cite{auslender2000lagrangian}, i.e.,
\begin{align*}
\left[
\begin{array}{c}
\nabla_{u_1} \mc L_1(u_1,u_2,\sigma_1,\sigma_2,\lambda_1,\lambda_2)\\
\nabla_{\sigma_1} \mc L_1(u_1,u_2,\sigma_1,\sigma_2,\lambda_1,\lambda_2)\\
\nabla_{u_2} \mc \mc L_2(u_1,u_2,\sigma_1,\sigma_2,\lambda_1,\lambda_2)\\
\nabla_{\sigma_2} \mc \mc L_2(u_1,u_2,\sigma_1,\sigma_2,\lambda_1,\lambda_2)\\
\sigma_1 - u_2\\
\sigma_2 - u_1
\end{array}
\right] =
%
\left[
\begin{array}{c}
2u_1 + \sigma_1 +1 - \lambda_2\\
u_1 +\lambda_1\\
2u_2 + \sigma_2 +1 - \lambda_1\\
u_2 +\lambda_2\\
\sigma_1 - u_2\\
\sigma_2 - u_1
\end{array}
\right] 
\overset{!}{
=}
%
 \mathbf{0},
\end{align*}
which yields v-GNEs of the form $(u_1^*, u_2^*) = (x,-1/2-x)$, with $x \in \mathbb R$.
{\hfill $\square$}
}

@@

\query[R4-5]{The objective function associated to the power traded with the main grid seems not so natural. If the electricity unit price, $c_{h}^{mg}$ in equation (9) is considered linear (instead of quadratic) with respect to the total consumption, this would simplify this term, further becoming (11) a decomposable function. Could the authors elaborate what is the advantage of using a quadratic term here? Is worth the improvement it could add with respect to the difficulty it brings to the model?}
\ans{By the formulation in \red{(9)--(11)}, which is based on \cite{atzeni2013}, the electricity price from the main grid follows a linear inverse demand function where it increases linearly as the total consumption increases. On the other hand, if  $c^{\mathrm{mg}}$ in (9) is a linear function, the electricity unit price is assumed to be constant, which is not the case in the practice. In our view, the former represent better the electricity unit price than the latter, as supported by \cite{atzeni2013noncooperative} and some references therein. In addition, as discussed in \cite[Section II-D]{atzeni2013noncooperative}, the cost function in \red{(11)} can represent either the actual energy cost (as a result of energy generation and transmission costs by the main grid) or simply a pricing function designed to incentivize load-shifting by the end users.}

\query[R4-6]{ Taking into account comment 1, suppose that we consider a reformulation of the problem to avoid coupling objective functions, and we employ the Alternating Direction Methods of Multipliers (ADMM) to solve the P2P market game in (1). Could the authors elaborate if this would be possible? Would the application of the ADDM also guarantees the convergence to a variational GNE of the P2P market game? What are the advantages of using Algorithm 1 proposed in the paper to solve the problem?  }
\ans{\green{Indeed, the only trivial case in which v-GNEs and social optima, i.e., the minimizers of the social welfare, coincide is for games with decoupled objective functions. This fact can be shown, for example, by a direct comparison of their KKT systems. In this special case, the original v-GNE seeking problem and the social welfare minimization problem are equivalent. Hence, the latter can be tackled via existing distributed optimization algorithms, e.g. ADMM. Unfortunately, in general, it is not possible to trivially reformulate games with coupled objectives as optimization problems -- see R4-3 for a simple counter-example in which a straightforward decoupling of the local objective functions fails to preserve the original equilibria. This simple fact motivated many authors in different research communities to design ad-hoc algorithms for games and generalized games with coupled cost functions, see e.g.  \cite{yi2019,paccagnan2019,liang2017distributed, koshal2016distributed,gadjov2020single,belgioioso2020b}.
}
}

@@

\query[R4-7]{ Regarding to the scalability of the algorithms, the author evaluates the convergence of the algorithm with respect to the number of agents and the connectivity, but they do not provide any information about the computational times. This is relevant to evaluate if it could be applied in a real large-scale system. }
\ans{We thank the Reviewer for raising this point. We agree that the computational time is relevant in practice and thus we have added the plots of the total computational time/agent in the scalability test section, %\red{particularly for the simulations with different numbers of agents} 
	(Figures \red{9--10} in the revised manuscript as also shown below). Since at each iteration, each agent only needs to update its own decision variable and some dual variables associated with the coupling constraints, we can see that the total computational time/agent is relatively low as the day-ahead problem is only solved once per day. \comment{show the new figures}
}

 \query[R4-8]{ Given a prosumer $i$, when Algorithm 1 is applied, in order to update the dual variable related to reciprocity constraints, the value of the power traded with neighbors plays an important role. In detail, I mean $\{p^{tr}_{(j,i)}\}_{j\in \mc N_{i}}$, the forward trade estimate to prosumer $i$. It seems that, depending on the order the prosumers are considered, the coordination of these variables could change, so that the convergence of the algorithm. Hence, it could be interesting to analyze how this affect to the behavior of the algorithm. For example, could it benefit to consider first prosumers with less neighbors, since maybe they reach faster an equilibrium? }
 	\ans{\green{
In the proposed market-clearing mechanism, all the local (primal and dual) updates of the prosumers are performed in parallel. We explicitly clarified this point in the new algorithm layout (see the new Algorithm 1).
 	}}

### 2022-01-11 papers/02 journal v2/journal_v14.tex
g

%\paragraph*{literature survey: what are currently studied}
Focusing on spot markets, i.e., day-ahead and intra-day markets, each prosumer has to decide its energy production and consumption over a certain time horizon, with the objective of minimizing its own economic cost while satisfying its physical and operational constraints. 
Most of existing works formulate such peer-to-peer (P2P) markets via game-theoretic or multi-agent optimization frameworks \cite{tushar2018,tushar2020,lecadre2020,cui2020,zhang2020,baroche2019,sorin2019}. For instance, the authors of \cite{tushar2018} provide a literature survey of early works on game-theoretic P2P market models. {More recently, \cite{tushar2020} considers a coalition game approach for peer-to-peer trading of prosumers with storage units.}  Furthermore, \cite{lecadre2020,cui2020,zhang2020,baroche2019,sorin2019} propose economic dispatch formulations where energy trading is incorporated as coupling (reciprocity) constraints and each prosumer has an objective that depends on local decision variables only.

Generalizing the previous papers, our preliminary work in  \cite{belgioioso2020energy} does not only consider multi-bilateral trading  but also trading with the main grid, which extends the coupling to both constraints and objective functions. 
%
Mathematically, clearing the resulting P2P market corresponds to finding a generalized Nash equilibrium (GNE), namely, a configuration in which no prosumer has an incentive to unilaterally deviate. 
%
Similarly, \cite{wang2021} formulates a generalized Nash game of energy sharing or a multilateral (instead of bilateral) trading among prosumers, and moreover, a distributed GNE seeking algorithm is designed to find a solution of the market equilibrium problem. 
%
In parallel, we note that operator-theoretic approaches have been effectively exploited to design distributed methods that solve GNE problems under the least restrictive assumptions \cite{paccagnan2019,belgioioso2020b,gadjov2020single,belgioioso2020semi,
	bianchi2020fast}.

%\paragraph*{highlight on considering physical constraints}
In practice, however, direct trading among prosumers might jeopardize system reliability, for which network operators are responsible. Therefore, when designing energy management mechanism for a distribution grid, one must also consider the role of network operators and the reliability of the system itself.
%
For example, \cite{qin2018,morstyn2019} treat decentralized markets and operational reliability separately, and propose market-clearing mechanisms where decentralized market solutions must be approved by a network operator based on the system operational constraints. 
%
An alternative is based on incorporating network charges, which may reflect utilization fees and network congestion, into the market formulation, as discussed in \cite{baroche2019b,paudel2020}. Differently, \cite{moret2020,zhang2020} include network operators as players in the market and impose operational requirements of the network as constraints in the market problem, which is formulated as a multi-agent optimization. Nevertheless, none of these works simultaneously consider coupled objectives and constraints, implying the inapplicability of their decentralized mechanisms to our market formulation.

@@

\begin{figure}
	\centering
	\includegraphics[scale=1.4]{figures/37bus_ex}
	\caption{A modified IEEE 37-bus network with 12 prosumers (boxes) and 15 passive loads (black triangles). Busses are represented by black circles, physical lines in $\mc L$ by solid lines, whereas trading relations ($\mc E$) by dash-dotted blue lines.
	}
	\label{fig:GGp}
\end{figure} 
%\subsubsection*{Types of market considered}
We focus on P2P spot markets, i.e., day-ahead and intra-day markets, similarly to \cite{sousa2019,lecadre2020,moret2020}. Thus, we denote the time horizon for which the decisions are computed by $\mc H = \{1,2,\dots,H \}$. For instance, in a day-ahead market, typically, the sampling period is one hour and the time horizon is $H=24$ hours. Moreover, as in \cite{moret2020}, we also include the physical constraints of the distribution network to ensure that a solution is not only economically optimal but also meets the standards of the DNO.

@@

\smallskip
We consider a linear approximation of power-flow equations, which is standard in the literature of P2P markets, e.g., \cite{yang2019,moret2020}. Specifically, for each bus $y \in \mc B$, it must hold that
\begin{equation}
\sum_{i \in \mc P_y^\mathrm{b}}p_i^{\mathrm{d}}+ \sum_{i\in \mc N_y^{\mathrm b}} \eta_i-p_{y}^{\mathrm{tg}}=\sum_{z\in \mc B_y}p_{(y,z)}^{\ell}, \label{eq:p_bal_p}
\end{equation}
where $\eta_i$ is the real power injection of prosumer $i$, i.e.,
\begin{equation}\label{eq:etai}
\eta_i:= p_i^{\mathrm d}-p_{i}^{\mathrm{di}}-\edit{p_{i}^{\mathrm{ds}}+p_{i}^{\mathrm{ch}}}
\end{equation}
Equation \eqref{eq:p_bal_p} models the local power balance of bus $y$, similarly to \eqref{eq:l_pow_b} although now it relates power generation, consumption, and line powers. Moreover, it must hold that
\begin{subequations}
\small
\label{eq:pf_pq}
\begin{align}
&p_{(y,z)}^{\ell} = B_{(y,z)}\left({\theta_y - \theta_z} \right) - G_{(y,z)}\left(v_y - v_z \right), \, \, \forall z \in \mc B_y, \label{eq:pf_p}\\
&q_{(y,z)}^{\ell} = G_{(y,z)}\left({\theta_y - \theta_z} \right) + B_{(y,z)}\left(v_y - v_z \right), \, \,  \forall z \in \mc B_y, \label{eq:pf_q}
\end{align}
\label{eq:pf_all}
\end{subequations}
which represent the power flow equations of line $(y,z)$ from the perspective of bus $y$, with $B_{(y,z)}$ and  $G_{(y,z)}$ denoting the susceptance and conductance, respectively, of line $(y,z)$. Note that by \eqref{eq:pf_p} and \eqref{eq:pf_q}, for each pair $(y,z) \in \mc L$, it holds that $p_{(y,z)}^{\ell} = {-p}_{(z,y)}^{\ell}$ and $q_{(y,z)}^{\ell} = {-q}_{(z,y)}^{\ell}$.

@@

\input{appendix}

>>>>

%\paragraph*{literature survey: what are currently studied}
Focusing on spot markets, i.e., day-ahead and intra-day markets, each prosumer has to decide its energy production and consumption over a certain time horizon, with the objective of minimizing its own economic cost while satisfying its physical and operational constraints. 
Most of existing works formulate such peer-to-peer (P2P) markets via game-theoretic or multi-agent optimization frameworks \cite{tushar2018,tushar2020,lecadre2020,cui2020,zhang2020,baroche2019,sorin2019}. For instance, the authors of \cite{tushar2018} provide a literature survey of early works on game-theoretic P2P market models. {More recently, \cite{tushar2020} considers a coalition game approach for peer-to-peer trading of prosumers with storage units.}  Furthermore, \cite{lecadre2020,cui2020,zhang2020,baroche2019,sorin2019} propose economic dispatch formulations where energy trading is incorporated as coupling (reciprocity) constraints and each prosumer has local decoupled objectives.

Generalizing the previous papers, our preliminary work in  \cite{belgioioso2020energy} does not only consider multi-bilateral trading  but also trading with the main grid, which extends the coupling to both constraints and objective functions. 
%
Mathematically, clearing the resulting P2P market corresponds to finding a generalized Nash equilibrium (GNE), namely, a configuration in which no prosumer has an incentive to unilaterally deviate. 
%
Similarly, \cite{wang2021} formulates a generalized Nash game of energy sharing or a multilateral (instead of bilateral) trading among prosumers, and proposes a distributed algorithm to find a solution of the market equilibrium problem. 
%
In parallel, we note that operator-theoretic approaches have been effectively exploited to design distributed methods that efficiently solve GNE problems under the least restrictive assumptions \cite{paccagnan2019,belgioioso2020b,gadjov2020single,belgioioso2020semi,
	bianchi2020fast}.

%\paragraph*{highlight on considering physical constraints}
In practice, however, direct trading among prosumers might jeopardize system reliability, for which network operators are responsible. Therefore, when designing energy management mechanisms for a distribution grid, one must also consider the role of network operators and the reliability of the system itself.
%
For example, \cite{qin2018,morstyn2019} treat decentralized markets and operational reliability separately, and propose market-clearing mechanisms where decentralized market solutions must be approved by a network operator based on the system operational constraints. 
%
An alternative is based on incorporating network charges, which may reflect utilization fees and network congestion, into the market formulation, as discussed in \cite{baroche2019b,paudel2020}. Differently, \cite{moret2020,zhang2020} include network operators as players in the market and impose network operational requirements as constraints in the market problem, which is formulated as a multi-agent optimization. Nevertheless, none of these works considers both coupled objectives and constraints.

@@

\begin{figure}
	\centering
	%\includegraphics[scale=1.4]{figures/37bus_ex}
	\includegraphics[width=\columnwidth]{figures/NF/37bus_sysIPE1}
	\caption{Left plot: A modified IEEE 37-bus network with 12 prosumers (boxes) and 15 passive loads (black triangles); busses are represented by black circles, physical lines in $\mc L$ by solid lines. Right plot: P2P trading network, whereas trading relations ($\mc E$) are represented by blue double-arrow lines.
	}
	\label{fig:GGp}
\end{figure}

%\begin{figure}
%	\centering
%	\includegraphics[scale=1.4]{figures/37bus_ex}
%	\caption{A modified IEEE 37-bus network with 12 prosumers (boxes) and 15 passive loads (black triangles). Busses are represented by black circles, physical lines in $\mc L$ by solid lines, whereas trading relations ($\mc E$) by dash-dotted blue lines.
%	}
%	\label{fig:GGp}
%\end{figure}

%\subsubsection*{Types of market considered}
We focus on P2P spot markets, i.e., day-ahead and intra-day markets, similarly to \cite{sousa2019,lecadre2020,moret2020}. Thus, we denote the time horizon for which the decisions are computed by $\mc H = \{1,2,\dots,H \}$. For instance, in a day-ahead market, typically, the sampling period is one hour and the time horizon is $H=24$ hours. Moreover, as in \cite{moret2020}, we also include the physical constraints of the distribution network to ensure that a solution is not only economically optimal but also meets the standards of the DNO.

@@

\smallskip
We consider a linear approximation of power-flow equations, which is standard in the literature of P2P markets, e.g., \cite{yang2019,moret2020}. Specifically, for each bus $y \in \mc B$, it must hold that
\begin{equation}
\sum_{i \in \mc P_y^\mathrm{b}}p_i^{\mathrm{d}}+ \sum_{i\in \mc N_y^{\mathrm b}} \eta_i-p_{y}^{\mathrm{tg}}=\sum_{z\in \mc B_y}p_{(y,z)}^{\ell}, \label{eq:p_bal_p}
\end{equation}
where $\eta_i$ is the active power injection of prosumer $i$, i.e.,
\begin{equation}\label{eq:etai}
\eta_i:= p_i^{\mathrm d}-p_{i}^{\mathrm{di}}-\edit{p_{i}^{\mathrm{ds}}+p_{i}^{\mathrm{ch}}}
\end{equation}
Equation \eqref{eq:p_bal_p} models the local power balance of bus $y$, similarly to \eqref{eq:l_pow_b} although now it relates power generation, consumption, and line powers. Moreover, it must hold that
\begin{subequations}
\small
\label{eq:pf_pq}
\begin{align}
&p_{(y,z)}^{\ell} = B_{(y,z)}\left({\theta_y - \theta_z} \right) - G_{(y,z)}\left(v_y - v_z \right), \, \, \forall z \in \mc B_y, \label{eq:pf_p}\\
&q_{(y,z)}^{\ell} = G_{(y,z)}\left({\theta_y - \theta_z} \right) + B_{(y,z)}\left(v_y - v_z \right), \, \,  \forall z \in \mc B_y, \label{eq:pf_q}
\end{align}
\label{eq:pf_all}
\end{subequations}
which represent the power flow equations of line $(y,z)$ from the perspective of bus $y$, with $B_{(y,z)}$ and  $G_{(y,z)}$ denoting the susceptance and conductance, respectively, of line $(y,z)$. Note that by \eqref{eq:pf_p} and \eqref{eq:pf_q}, for each pair $(y,z) \in \mc L$, it holds that $p_{(y,z)}^{\ell} = {-p}_{(z,y)}^{\ell}$ and $q_{(y,z)}^{\ell} = {-q}_{(z,y)}^{\ell}$.

@@

\newpage
\input{appendix}
